---
author: "Eleanor Fewings"
output: html_document
---

# start_section  
```{r start_section}

# Time stamp
Sys.time()

# Clean up
rm(list=ls())

# Load libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(Seurat)
library(stringr)
library(progeny)
library(tibble)
library(pheatmap)
library(stringr)
library(knitr)
library(DoMultiBarHeatmap)
library(fgsea)
library(data.table)
library(EnhancedVolcano)
library(ggpubr)

# Folders
setwd("C:/Users/ellie/OneDrive/Saez/Sebastiaan_BK/")
opts_knit$set(root.dir = 'C:/Users/ellie/OneDrive/Saez/Sebastiaan_BK/')

# Source functions
source("C:/Users/ellie/OneDrive/Saez/Sebastiaan_BK/scripts/source.R")


```

## Load data
```{r load}

#Load data
data <- readRDS("CK226.rds")

#Separate metadata for exploration
metadata <- data@meta.data

#Read viraltrack information
vt <- read.table("viraltrack/AllFilteredBarcodes/Expression_table.tsv", sep="\t", stringsAsFactors = FALSE, header=TRUE, row.names=1) %>% t()

#Summarise counts of viraltrack
vtsum <- rowSums(vt[,1:3]) %>% as.data.frame()

#Add a "-1" to barcodes to match cellranger barcodes
row.names(vtsum) <- paste(row.names(vtsum), "-1", sep="")

#Rename column
colnames(vtsum) <- "VTsum"

## Summary statistics
#Maximum viral reads
max(vtsum$.)

#Number of cells with no viral reads
sum(vtsum$. == 0)

#Number of cells with viral reads
sum(vtsum$. != 0)

#Proportion of cells with viral reads
sum(vtsum$. != 0)/nrow(vtsum)

#Mean number of viral reads in cells with viral reads
mean(vtsum[vtsum$. != 0,])

#Merging vtsum to metadata
metadata.vt <- merge(metadata, vtsum, by="row.names", all.x=TRUE)

#Count number of cells with viral reads, without viral reads and those that weren't detected
table(metadata.vt$VTsum, exclude = NULL)

#Create binary
metadata.vt$VTbin <- ifelse(is.na(metadata.vt$VTsum) | metadata.vt$VTsum == "0" , "No viral reads", "Viral reads")

#Add rownames back (removed by merge)
row.names(metadata.vt) <- metadata.vt$Row.names

#Convert NA to 0
metadata.vt$VTsum[is.na(metadata.vt$VTsum)] <- 0

#Check rownames match original data order
sum(row.names(metadata) != row.names(metadata.vt))

#Add viral track data to seurat object
data$VTsum <- metadata.vt$VTsum
data$VTbin <- metadata.vt$VTbin

#Cleanup
rm(metadata.vt, vt, vtsum)

```

#Find background viral level
```{r raw, eval = FALSE, echo = FALSE}

#Read viraltrack information
raw <- read.table("viraltrack/AllRawBarcodes/Expression_table.tsv", sep="\t", stringsAsFactors = FALSE, header=TRUE, row.names=1) %>% t()

#Add a "-1" to barcodes to match cellranger barcodes
row.names(raw) <- paste(row.names(raw), "-1", sep="")

#Count raw cells
nrow(raw)

#Remove cells that are in metadata
raw.filt <- raw[!row.names(raw) %in% row.names(metadata),]

nrow(raw.filt)

rawsum <- rowSums(raw.filt[,1:4]) %>% as.data.frame()

## Summary statistics
#Maximum viral reads
max(rawsum$.)

#Number of cells with no viral reads
sum(rawsum$. == 0)

#Number of cells with viral reads
sum(rawsum$. != 0)

#Proportion of cells with viral reads
sum(rawsum$. != 0)/nrow(rawsum)

#Mean number of viral reads in cells with viral reads
mean(rawsum[rawsum$. != 0,])


```



#Relabel clusters
```{r label}

#Create list of new cluster labels
label <- list("0" = "T-Lymphocytes",
          "1"="T-Lymphocytes",
          "2"="T-Lymphocytes",
          "3"="B-Lymphocytes",
          "4"="Endothelial cells",
          "5"="Proximal tubular epithelium",
          "6"="Macrophages",
          "7"="NK-cells",
          "8"="Distal tubular epithelium",
          "9"="Fibroblasts/Pericytes",
          "10"="Endothelial cells (glomerular)",
          "11"="Plasmacells",
          "12"="Intercalated cells/Collecting duct",
          "13"="Podocytes")

#Rename clusters
data <- RenameIdents(data, label)

#Store idents in metadata
data$celltype <- Idents(data)

#Plot
DimPlot(data, reduction = "umap", label = TRUE)

#Change identity to viral count
Idents(data) <- "VTbin"

#Plot
DimPlot(data, reduction = "umap", label = FALSE, pt.size = 2, order = "Viral reads")


```

#Count proportions of infected and uninfected cells per cell type
```{r prop}

#recrete metadata
metadata <- data@meta.data

#Create count
sum.vt <- metadata %>% add_count(seurat_clusters, VTbin)

#Create percentage
sum.vt <- sum.vt %>% group_by(seurat_clusters) %>% mutate(perc=n/length(seurat_clusters))

#Create cluster-celltype column
sum.vt$clusttype <- paste(sum.vt$seurat_clusters, sum.vt$celltype, sep="-")

#Shorten and make unique
sum.vt <- sum.vt %>% subset(select=c(VTbin, clusttype, perc, n)) %>% unique()

#Plot bar graph (number infected)
ggplot(sum.vt, aes(x=VTbin, y=n, fill=VTbin)) +
  geom_bar(stat="identity") +
  facet_wrap(~clusttype)

#(Percentage infected)
ggplot(sum.vt, aes(x=VTbin, y=perc, fill=VTbin)) +
  geom_bar(stat="identity") +
  facet_wrap(~clusttype)


```


#compare viral affected cells per cluster
```{r de}

#Compare markers per cluster
response.0 <- subset(data, subset= seurat_clusters == 0) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.1 <- subset(data, subset= seurat_clusters == 1) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.2 <- subset(data, subset= seurat_clusters == 2) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.3 <- subset(data, subset= seurat_clusters == 3) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.4 <- subset(data, subset= seurat_clusters == 4) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.5 <- subset(data, subset= seurat_clusters == 5) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.6 <- subset(data, subset= seurat_clusters == 6) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.7 <- subset(data, subset= seurat_clusters == 7) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.8 <- subset(data, subset= seurat_clusters == 8) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
#response.9 <- subset(data, subset= seurat_clusters == 9) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
#response.10 <- subset(data, subset= seurat_clusters == 10) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
#response.11 <- subset(data, subset= seurat_clusters == 11) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
response.12 <- subset(data, subset= seurat_clusters == 12) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)
#response.13 <- subset(data, subset= seurat_clusters == 13) %>% FindMarkers(ident.1 = "Viral reads", ident.2 = "No viral reads", verbose = FALSE)

#Add cluster label
response.0$cluster <- "0"
response.1$cluster <- "1"
response.2$cluster <- "2"
response.3$cluster <- "3"
response.4$cluster <- "4"
response.5$cluster <- "5"
response.6$cluster <- "6"
response.7$cluster <- "7"
response.8$cluster <- "8"
#response.9$cluster <- "9"
#response.10$cluster <- "10"
#response.11$cluster <- "11"
response.12$cluster <- "12"
#response.13$cluster <- "13"

#Add gene column
response.0$gene <- row.names(response.0)
response.1$gene <- row.names(response.1)
response.2$gene <- row.names(response.2)
response.3$gene <- row.names(response.3)
response.4$gene <- row.names(response.4)
response.5$gene <- row.names(response.5)
response.6$gene <- row.names(response.6)
response.7$gene <- row.names(response.7)
response.8$gene <- row.names(response.8)
#response.9$gene <- row.names(response.9)
#response.10$gene <- row.names(response.10)
#response.11$gene <- row.names(response.11)
response.12$gene <- row.names(response.12)
#response.13$gene <- row.names(response.13)

#Merge back results
markers <- rbind(response.0, response.1, response.2, response.3, response.4, response.5, response.6, response.7, response.8, response.12)

#Make volcano plot of all responses
EnhancedVolcano(markers,
    lab = markers$gene,
    x = 'avg_log2FC',
    y = 'p_val_adj')
  
#Clean up
#rm(response.0, response.1, response.2, response.3, response.4, response.5, response.6, response.7, response.8, response.12)

#select top 10 per cluster (and select significant only)
top <- markers %>% filter(p_val_adj < 0.05) %>% group_by(cluster) %>% top_n(-8, p_val_adj) %>% top_n(8, avg_log2FC)

#Plot
DoMultiBarHeatmap(object = data, features = top$gene, group.by = "VTbin", additional.group.by = "seurat_clusters")

#Write response files and per cluster volcanos
responses <- list(response.0=response.0, 
                  response.1=response.1, 
                  response.2=response.2, 
                  response.3=response.3, 
                  response.4=response.4, 
                  response.5=response.5, 
                  response.6=response.6, 
                  response.7=response.7, 
                  response.8=response.8, 
                  response.12=response.12)

for (i in 1:length(responses)) {
  #Select response file
  response <- responses[i] %>% as.data.frame()
  #Edit colnames
  colnames(response) <- colnames(response) %>% sub(paste(names(responses)[i], ".", sep=""), "", .)
  #Get cluster number
  clust <- names(responses)[i] %>% sub("response.", "", .)
  #Find celltype
  type <- label[clust] %>% unlist() %>% sub("/", "_", .) %>% sub(" ", "_", .)
  #Set csv file name
  file <- paste("markergenes/cluster", clust, "_", type, ".csv", sep="")
  #Write csv
  #write.csv(response, file)
  #Set pdf filename
  pdfn <- paste("volcano/cluster", clust, "_", type, "volcano.pdf", sep="")
  #Write volcano
  pdf(file = pdfn, width = 10, height = 9)
  print(EnhancedVolcano(response,
    lab = response$gene,
    x = 'avg_log2FC',
    y = 'p_val_adj'))
  dev.off()
  }

```

#create heatmap of specific clusters
```{r heatmap}

#clusters to look at
#0 and 1 - T Lymphocytes
#11 - plasma cells
#12 - intercalated cells
#13 - podocytes
#4 - endothelial
#5 - proximal tubular epithelial cells
#8 - distal tubular epithelial

#Filter summary to find percentage of each celltype containing viral reads
viral <- sum.vt %>% filter(VTbin == "Viral reads")

#Find upper quartile
quarts <- summary(viral$perc)[5]

#Show clusters in upper quartile
viral %>% filter(perc >= quarts)

#Set cc as identity
Idents(data) <- "VTbin"

#Set cluster as metadata
data$cluster <- data@meta.data$seurat_clusters

#Cluster 0
c0 <- subset(data, subset= cluster == 0)

top.0 <- response.0[order(response.0$p_val_adj, response.0$avg_log2FC),][1:60,]

DoMultiBarHeatmap(object = c0, features = top.0$gene, group.by = "VTbin", additional.group.by = "celltype")

#Cluster 1
c1 <- subset(data, subset= cluster == 1)

top.1 <- response.1[order(response.1$p_val_adj, response.1$avg_log2FC),][1:60,]

DoMultiBarHeatmap(object = c1, features = top.1$gene, group.by = "VTbin", additional.group.by = "celltype")


#Cluster 4
c4 <- subset(data, subset= cluster == 4)

top.4 <- response.4[order(response.4$p_val_adj, response.4$avg_log2FC),][1:60,]

DoMultiBarHeatmap(object = c4, features = top.4$gene, group.by = "VTbin", additional.group.by = "celltype")


#Cluster 5
c5 <- subset(data, subset= cluster == 5)

top.5 <- response.5[order(response.5$p_val_adj, response.5$avg_log2FC),][1:60,]

DoMultiBarHeatmap(object = c5, features = top.5$gene, group.by = "VTbin", additional.group.by = "celltype")

#Cluster 8
c6 <- subset(data, subset= cluster == 6)

top.6 <- response.6[order(response.6$p_val_adj, response.6$avg_log2FC),][1:60,]

DoMultiBarHeatmap(object = c6, features = top.6$gene, group.by = "VTbin", additional.group.by = "celltype")


#Cluster 8
c8 <- subset(data, subset= cluster == 8)

top.8 <- response.8[order(response.8$p_val_adj, response.8$avg_log2FC),][1:60,]

DoMultiBarHeatmap(object = c8, features = top.8$gene, group.by = "VTbin", additional.group.by = "celltype")

#Cluster 12
c12 <- subset(data, subset= cluster == 12)

top.12 <- response.12[order(response.12$p_val_adj, response.12$avg_log2FC),][1:60,]

DoMultiBarHeatmap(object = c12, features = top.12$gene, group.by = "VTbin", additional.group.by = "celltype")


```

#From Sebastiaan's email
## Proximal tubular epithelial cells
Infected cells should theoretically show:
*TP53 increase
Tap73 increase
DNMT1 increase
*CDKN1A decrease
COLIVA1 decrease

Senescent cells (or cells in a late stage of viral infection) should theoretically show:
*CDKN1A increase 
CDKN2A increase
GLS1 increase
E2F7 increase
LMNB1 increase

#Check downstream expression for other cluster
```{r geneexpr_sets}

g.list <- c("TP53", "Tap73", "DNMT1", "CDKN1A", "COLIVA1", "CDKN2A", "GLS1", "E2F7", "LMNB1")

#Create scatter plots of expression in each gene for different clusters
##C5
expr.plot(5, g.list, pval=FALSE)

##C8
expr.plot(8, g.list, pval=FALSE)

##C12
expr.plot(12, g.list, pval=FALSE)

#senescent genes
senescent.g <- c("SIRT1",
"BCL2",
"AKT1",
"VEGFA",
"MCM7",
"MTHFD2",
"CDKN2B",
"CCND1",
"HDAC1",
"E2F1",
"E2F3",
"CCNE1",
"JUN",
"PCNA",
"SMAD3",
"MAPK1",
"SIRT6",
"TGFB2",
"SMAD1",
"CDKN2C",
"CCNE2",
"TP63",
"PRKDC",
"RBL1",
"MPAK3"
)

#Create scatter plots of expression in each gene for different clusters
##C5
expr.plot(5, senescent.g)

##C8
expr.plot(8, senescent.g)

##C12
expr.plot(12, senescent.g)

#Create senescence score for all clusters
score <- AddModuleScore(
  object = data,
  features = list(senescent.g),
  ctrl = 5,
  name = 'senes'
)

#Subset metadata with senescence scores
score.meta <- score@meta.data

#Plot histogram of senescence results
ggplot(score.meta, aes(x=senes1)) +
  geom_histogram()

#Plot histogram of senes results colouring by infection with BK
ggplot(score.meta, aes(x=senes1, fill=VTbin)) +
  geom_histogram()

#Plot histogram of senes results colouring by infection with BK
ggplot(score.meta, aes(x=VTbin, y=senes1, fill=VTbin)) +
  geom_boxplot() +
  stat_compare_means(method = "t.test", label.y = 0.6) +
  ggtitle("Comparison of senescence scores between virally infected cells and controls")

#Compare senescence scores per cluster (writes to pdf)
score.comparison(c5)

score.comparison(c8)

score.comparison(c12)

```

#Run GSEA
```{r gsea}

#Load genesets
p53 <- read.table("genesets/p53.txt", sep="\t", skip=2, stringsAsFactors = FALSE, header=FALSE)
g2m <- read.table("genesets/g2m.txt", sep="\t", skip=2, stringsAsFactors = FALSE, header=FALSE)
pi3k <- read.table("genesets/pi3k.txt", sep="\t", skip=2, stringsAsFactors = FALSE, header=FALSE)
tnfa <- read.table("genesets/tnfa.txt", sep="\t", skip=2, stringsAsFactors = FALSE, header=FALSE)

#Cluster 5

#p53
p53.5 <- response.5[row.names(response.5) %in% p53$V1 & response.5$p_val_adj < 0.05,]
DoMultiBarHeatmap(object = c5, features = p53$V1, group.by = "VTbin", additional.group.by = "celltype")

#g2m
g2m.5 <- response.5[row.names(response.5) %in% g2m$V1 & response.5$p_val_adj < 0.05,]
DoMultiBarHeatmap(object = c5, features = g2m$V1, group.by = "VTbin", additional.group.by = "celltype")

#pi3k
pi3k.5 <- response.5[row.names(response.5) %in% pi3k$V1 & response.5$p_val_adj < 0.05,]
DoMultiBarHeatmap(object = c5, features = pi3k$V1, group.by = "VTbin", additional.group.by = "celltype")

#g2m
tnfa.5 <- response.5[row.names(response.5) %in% tnfa$V1 & response.5$p_val_adj < 0.05,]
DoMultiBarHeatmap(object = c5, features = tnfa$V1, group.by = "VTbin", additional.group.by = "celltype")



```

#GSEA averaging set
```{r gsea_avg}

#Plot average expression over gene sets
gsea.expr.plot(5, p53$V1, "VTbin")
gsea.expr.plot(5, g2m$V1, "VTbin")
gsea.expr.plot(5, pi3k$V1, "VTbin")
gsea.expr.plot(5, tnfa$V1, "VTbin")

```

#Run FGSEA
```{r fgsea}

#Select genes to run analysis
bg.genes <- row.names(data)

#DE clusters
de.files <- list(response.0, response.1, response.2, response.3, response.4, response.5, response.6, response.7, response.8, response.12)

for (file in de.files){
  markers <- file[order(file$avg_log2FC, decreasing = TRUE),]
  rownames(markers) <- markers$gene
  
  # Get cell type
  clust <- unique(markers$cluster)
  cell.type <- label[names(label) == clust] %>% unlist() %>% str_replace("/", "_") %>% str_replace(" ", "_")
  cell.type <- paste(cell.type, paste0("cluster", clust), sep="_")
  outdir <- "GSEA"
  dir.create(outdir)
  
  # GSEA
  stats <- markers$avg_log2FC
  names(stats) <- rownames(markers)
  stats <- stats[!is.na(stats)]
  file.prefix <- paste('viral_vs_nonviral_', cell.type, sep="") 
  
  # GO
  run_gsea(bg.genes = bg.genes, stats = stats, 
           category = 'C5', subcategory = 'BP',
           out.dir = outdir, plot.title = 'GO',
           file.prefix = file.prefix)
  
  # PID
  run_gsea(bg.genes = bg.genes, stats = stats, 
           category = 'C2', subcategory = 'PID',
           out.dir = outdir, plot.title = 'PID',
           file.prefix = file.prefix)
  
  # Immunological signature
  run_gsea(bg.genes = bg.genes, stats = stats, 
           category = 'C7',
           out.dir = outdir, plot.title = 'Immunological Signature',
           file.prefix = file.prefix)
}



```

